---
title: "Aides Drug"
author: "Tanya Claros"
date: "Summer 2022"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
Females = filter(triUmp,gender=="F")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that sub population with the filter command.
  
$H_0$: The drug was not effective in preventing females from contracting the virus.   
$H_A$: The drug was effective in preventing females from contracting the virus.  
  
# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
  
Since both of our variables are categorical, we will use the bar plot with a chisq.test.    

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=Females)
barchartGC(~infected + treatment,data=Females, type="percent")

```
  
The drug appears to make females more susceptible to the virus. 
  
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
  
```{r}
table1 <- xtabs(~infected + treatment, data=Females)
rowPerc(table1)
colPerc(table1)
```  
  
The table shows that if a female were to get the vaccine, she would be 1.4x more likely to contract the virus in comparison to the placebo group.   
    
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
  
```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```  
  
We reject the null hypothesis because the p-value is less than 0.05. Our p-value is 0.00007686. The fisher exact table is also significant in showing that we reject the null hypothesis because the p-value is 0.00006596.  
    
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
  
The original question was: Is the drug seem effective in preventing females from contracting aides.  $H_0$: The drug was not effective in preventing females from contracting the virus. $H_A$: The drug was effective in preventing females from contracting the virus. The table showed that females were 1.4x more susceptible in contracting aides with the preventative measures compared to the placebo. Our p-value is 0.00007686, smaller than 0.05 and because of that we reject the null hypothesis. 
  